---
title: "Import Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Import Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette describes how to import single-cell CRISPR screen data into `sceptre`. Users either can import data from the output of an existing software package (e.g., 10X CellRanger, the Parse Biosciences count matrix generation program) or from a set of R objects. We recommend that users import data from the output of an existing software package, as doing so decreases the probability of error.

## Import data from an existing software package

`sceptre` links up with two programs for upstream processing of single-cell CRISPR screen data: 10X Cellranger and the analogous software tool developed by Parse Biosciences. (We refer to the latter program as the "Parse program.") We describe how to import data from each of these tools below.

### Import from 10X CellRanger

The Get Started vignette (`vignette("sceptre")`) describes how to import data from 10X CellRanger into `sceptre`. We provide a minimal working example here:

```{r}

```

### Import from the Parse program

A function to import data from the Parse program is coming soon.

### Add extra covariates to a `sceptre_object`

```{r}
# simulate the covariate cell_type
cell_type <- sample(x = paste0("type_", 1:3),
                    size = 10000,
                    replace = TRUE) |> factor()
extra_covariates <- data.frame(cell_type = cell_type)

# add extra covariate to sceptre_object
sceptre_object <- add_extra_covariates(sceptre_object,
                                       extra_covariates = extra_covariates)
```

## Import data from R objects

